\documentclass[14pt,twoside]{report}

\usepackage{amsfonts,amsmath,amssymb,stmaryrd,indentfirst}
\usepackage{epsfig,graphicx,times,psfrag}
\usepackage{natbib}
\usepackage{pdfpages,enumerate}% ,enumitem}
 
%%% Rotating Page
\usepackage{pdflscape}
\usepackage{afterpage}
%\usepackage{capt-of}% or use the larger `caption` package
%\usepackage{lipsum}% dummy text


\usepackage[pdftex,bookmarks,colorlinks=true,urlcolor=blue,citecolor=blue]{hyperref}

%\usepackage{fancyhdr} %%%%
%\pagestyle{fancy}%%%%
\pagestyle{empty}
\def\newblock{\hskip .11em plus .33em minus .07em}

\setlength\textwidth      {16.5cm}
\setlength\textheight     {22.0cm}
\setlength\oddsidemargin  {-0.3cm}
\setlength\evensidemargin {-0.3cm}

\setlength\headheight{0in} 
\setlength\topmargin{0.cm}
\setlength\headsep{1.cm}
\setlength\footskip{1.cm}
\setlength\parskip{0pt}

%%%
%%% space between lines
%%%
\renewcommand {\baselinestretch}{1.5}

\begin{document}

Make sure that you proofread the text before resubmitting. In the current report, there are a large number of typos and unrevised sentences. Several sentences are confusing and disconnected with no clear objectives and inter-connectives.

\begin{enumerate}[A.]
%
   \item Format and Style:
        \begin{enumerate}[i)]
           \item The report {\bf must} contain an \underline{abstract} and  \underline{aims $\&$ objectives} section;
           \item {\bf All} sections {\bf must} be numbered;
           \item Equations should be numbered consecutively, with the number set flush right and enclosed in parentheses either either as single numbers (e.g., (5)) or section-based (e.g., (5.1));
           \item References: 
              \begin{itemize}
                  \item Must not contain 'et al.';
                  \item Remove `In' from paper references. `In' is used for book chapters;
              \end{itemize} 
           \item Second-year report and thesis are formal documents, in which you demonstrate a high level of expertise with respect to a (few) topic(s) of your field study. As such, every argument should be in the appropriate place and corroborated by data (produced by yourself) and/or by the literature. As a formal document, you should use formal and concise language (and therefore avoiding colloquial expressions such as, `generally speaking');
        \end{enumerate}
%
   \item The resubmitted report must include:
       \begin{enumerate}[i)]
           \item Main aim of the research work;
           \item List of objectives; 
           \item Expected deliverables and milestones;
           \item List of tasks to be undertaken with expected time for completion.
       \end{enumerate}
%
   \item A few technical comments about the report: 
       \begin{enumerate}[i)]
          \item Reproducibility of (numerical) experiments is a key aspect in scientific communication. All results should be accompanied by a full description of the numerical experiment;
          \item Section `Time-stepping method': The text is very confusing and need to be rewritten. Arguments need to be fully explained (with associated references). For example, Eqns. 3-5 describe the new integrator with no reasoning for the associated error (Eqn. 5);
          \item You should define 3 key concepts: stability, convergence and accuracy;
          \item Section 'Event-driven method': Same comment from the previous section.  Also it is not clear is you are referring to total energy or dissipative energy (page 6);
          \item Most of the discussion on EDM's effectiveness (Eqns. 6-13 and Fig. 3) is based on the equation placed in Figure 1. This equation should be located (and explained) in the main text; 
          \item How do you plan to optimise / assess EDM ? 
       \end{enumerate}
%
\end{enumerate}
   
\end{document}
